---
title: "Pre-trial jail inmates in Connecticut analysis"
author: "Andrew Ba Tran"
date: "January 9, 2017"
output: html_document
---

# Some charts
 *[x] Fix annotation locations
 *[x] Start total numbers from zero
 *[x] Get per capita based on race via Census
 *[x] Histogram of bail amounts
 *[ ] Map of facilities with circles on how many there are
 *[ ] Chart out bond amount by race
 *[ ] Correlation between bail and length of time in jail
 *[ ] Visualize the comingss and goings and those left behind...
 
 
 
```{r libraries, warning=F, message=F}
library(dplyr)
library(tidyr)
library(ggplot2)
library(lubridate)
library(knitr)
library(ggalt)
library(extrafont)
library(grid)
library(gridExtra)
library(stringr)
library(DT)
```

```{r data_ingesting, warning=FALSE}
update <- read.csv("https://data.ct.gov/api/views/b674-jy6w/rows.csv")
```


```{r race_date_total, fig.width=9, fig.height=6}
# by date, race total
race_total_date <- update %>%
  group_by(DOWNLOAD.DATE, RACE) %>%
  summarise(total=n()) %>%
  filter(RACE=="BLACK" | RACE=="HISPANIC" | RACE=="WHITE")

race_total_date$DOWNLOAD.DATE <- mdy(race_total_date$DOWNLOAD.DATE)


black_last <- race_total_date %>%
  filter(RACE=="BLACK") 
black_last$DOWNLOAD.DATE <-ymd(black_last$DOWNLOAD.DATE)
black_last <- arrange(black_last, desc(DOWNLOAD.DATE))
black_last <- black_last$total[1]


white_last <- race_total_date %>%
  filter(RACE=="WHITE") 
white_last$DOWNLOAD.DATE <-ymd(white_last$DOWNLOAD.DATE)
white_last <- arrange(white_last, desc(DOWNLOAD.DATE))
white_last <- white_last$total[1]

hispanic_last <- race_total_date %>%
  filter(RACE=="HISPANIC") 
hispanic_last$DOWNLOAD.DATE <-ymd(hispanic_last$DOWNLOAD.DATE)
hispanic_last <- arrange(hispanic_last, desc(DOWNLOAD.DATE))
hispanic_last <- hispanic_last$total[1]


gg <- ggplot(race_total_date, aes(x=DOWNLOAD.DATE, y=total, group=RACE, color=RACE)) + geom_line()
gg <- gg + labs(x=NULL, y="Inmates", title=paste("Accused pre-trial inmates in Connecticut jails"),
                caption="SOURCE: Department of Corrections\nAndrew Ba Tran/TrendCT.org")
gg <- gg + theme_bw(base_family="Calibri")
#gg <- gg + theme(panel.grid.major=element_blank())
#gg <- gg + theme(panel.grid.minor=element_blank())
gg <- gg + theme(panel.border=element_blank())
gg <- gg + theme(text = element_text(size=10))
#gg <- gg + theme(axis.ticks=element_blank())
#gg <- gg + theme(axis.text.x=element_blank())
gg <- gg + theme(plot.title=element_text(face="bold", family="Lato Black", size=22))
gg <- gg + theme(plot.subtitle=element_text(face="italic", size=9, margin=margin(b=12)))
gg <- gg + theme(plot.caption=element_text(size=12, margin=margin(t=10, r=80), color="#7a7d7e"))
gg <- gg + geom_label(data=race_total_date, aes(x=max(ymd(DOWNLOAD.DATE))+2, y=black_last, label=paste0("Black (", black_last,")"), hjust=0),
                      family="Helvetica", lineheight=0.95,
                      size=4.5, label.size=0, color="black")
gg <- gg + geom_label(data=race_total_date, aes(x=max(ymd(DOWNLOAD.DATE))+2, y=hispanic_last, label=paste0("Hispanic (", hispanic_last,")"), hjust=0),
                     family="Helvetica", lineheight=0.95,
                      size=4.5, label.size=0, color="black")
gg <- gg + geom_label(data=race_total_date, aes(x=max(ymd(DOWNLOAD.DATE))+2, y=white_last, label=paste0("White (", white_last,")"), hjust=0),
                      family="Helvetica", lineheight=0.95,
                      size=4.5, label.size=0, color="black")
gg <- gg + theme(legend.position="none")
gg <- gg + theme(plot.margin = unit(c(1, 10, 1, 1), "lines"))
gg

gb <- ggplot_build(gg)
gt <- ggplot_gtable(gb)

gt$layout$clip[gt$layout$name=="panel"] <- "off"
grid.draw(gt)
# ggsave(gt, file = "race_totals_time.png", width = 8, height = 6, type = "cairo-png")

## Ok now from zero

black_first <- filter(race_total_date, RACE=="BLACK" & DOWNLOAD.DATE=="2016-07-01")
black_first <- black_first$total
white_first <- filter(race_total_date, RACE=="WHITE" & DOWNLOAD.DATE=="2016-07-01")
white_first <- white_first$total
hispanic_first <- filter(race_total_date, RACE=="HISPANIC" & DOWNLOAD.DATE=="2016-07-01")
hispanic_first <- hispanic_first$total


race_total_date_adjusted <- race_total_date
race_total_date_adjusted$total <- ifelse(race_total_date_adjusted$RACE=="BLACK", race_total_date_adjusted$total-black_first, race_total_date_adjusted$total)
race_total_date_adjusted$total <- ifelse(race_total_date_adjusted$RACE=="HISPANIC", race_total_date_adjusted$total-hispanic_first, race_total_date_adjusted$total)
race_total_date_adjusted$total <- ifelse(race_total_date_adjusted$RACE=="WHITE", race_total_date_adjusted$total-white_first, race_total_date_adjusted$total)


black_last_adj <- race_total_date_adjusted %>%
  filter(RACE=="BLACK") 
black_last_adj$DOWNLOAD.DATE <-ymd(black_last_adj$DOWNLOAD.DATE)
black_last_adj <- arrange(black_last_adj, desc(DOWNLOAD.DATE))
black_last_adj <- black_last_adj$total[1]


white_last_adj <- race_total_date_adjusted %>%
  filter(RACE=="WHITE") 
white_last_adj$DOWNLOAD.DATE <-ymd(white_last_adj$DOWNLOAD.DATE)
white_last_adj <- arrange(white_last_adj, desc(DOWNLOAD.DATE))
white_last_adj <- white_last_adj$total[1]

hispanic_last_adj <- race_total_date_adjusted %>%
  filter(RACE=="HISPANIC") 
hispanic_last_adj$DOWNLOAD.DATE <-ymd(hispanic_last_adj$DOWNLOAD.DATE)
hispanic_last_adj <- arrange(hispanic_last_adj, desc(DOWNLOAD.DATE))
hispanic_last_adj <- hispanic_last_adj$total[1]

gg <- ggplot(race_total_date_adjusted, aes(x=DOWNLOAD.DATE, y=total, group=RACE, color=RACE)) + geom_line()
gg <- gg + labs(x=NULL, y="Inmates", title=paste("Accused pre-trial inmates in Connecticut jails since July 2016"),
                caption="SOURCE: Department of Corrections\nAndrew Ba Tran/TrendCT.org")
gg <- gg + theme_bw(base_family="Calibri")
gg <- gg + theme(panel.border=element_blank())
gg <- gg + theme(text = element_text(size=10))
gg <- gg + theme(plot.title=element_text(face="bold", family="Lato Black", size=22))
gg <- gg + theme(plot.subtitle=element_text(face="italic", size=9, margin=margin(b=12)))
gg <- gg + theme(plot.caption=element_text(size=12, margin=margin(t=10, r=80), color="#7a7d7e"))
gg <- gg + geom_label(data=race_total_date_adjusted, aes(x=max(ymd(DOWNLOAD.DATE))+2, y=black_last_adj, label=paste0("Black (", black_last_adj, ")"), hjust=0),
                      family="Helvetica", lineheight=0.95,
                      size=4.5, label.size=0, color="black")
gg <- gg + geom_label(data=race_total_date_adjusted, aes(x=max(ymd(DOWNLOAD.DATE))+2, y=hispanic_last_adj, label=paste0("Hispanic (", hispanic_last_adj, ")"), hjust=0),
                      family="Helvetica", lineheight=0.95,
                      size=4.5, label.size=0, color="black")
gg <- gg + geom_label(data=race_total_date_adjusted, aes(x=max(ymd(DOWNLOAD.DATE))+2, y=white_last_adj, label=paste0("White (", white_last_adj, ")"), hjust=0),
                      family="Helvetica", lineheight=0.95,
                      size=4.5, label.size=0, color="black")
gg <- gg + theme(legend.position="none")
gg <- gg + theme(plot.margin = unit(c(1, 10, 1, 1), "lines"))
gg

gb <- ggplot_build(gg)
gt <- ggplot_gtable(gb)

gt$layout$clip[gt$layout$name=="panel"] <- "off"
grid.draw(gt)

# NEXT

race_total_date$RACE <- gsub(" ", ".", race_total_date$RACE)

race_total_date <- update %>%
  group_by(DOWNLOAD.DATE, RACE) %>%
  summarise(total=n()) %>%
  spread(RACE, total) %>%
  mutate(total = `AMER IND` + ASIAN + BLACK + HISPANIC + WHITE) %>%
  mutate(American.Indian.per = round(`AMER IND`/total*100,2), Asian.per = round(ASIAN/total*100,2), Black.per = round(BLACK/total*100,2), Hispanic.per = round(HISPANIC/total*100,2), White.per = round(WHITE/total*100,2)) %>%
  select(DOWNLOAD.DATE, American.Indian.per, Asian.per, Black.per, Hispanic.per, White.per) %>%
  gather(Race, Percent, 2:6)

race_total_date$Race <- gsub(".per", "", race_total_date$Race)
race_total_date <- filter(race_total_date, Race!="American.Indian")

# Getting Census stats
library("censusapi")
source("keys.R")

# Black 
# B02001_003E
# White
# B02001_002E
# Hispanic

# Asian
# B02001_005E
# Total
# B02001_001E
#B03001_001E
#B03002_001E
#Not hispanic, white
# Total B03002_001E
# White B03002_003E
# Black B03002_004E
# Asian B03002_006E
# Hispanic B03002_012E




race2015 <- getCensus(name="acs5", 
    vintage=2015,
    key=census_key, 
    vars=c("NAME", "B03002_001E", "B03002_003E",
    "B03002_004E", "B03002_006E", "B03002_012E"), 
    region="state:*")

race2015 <- filter(race2015, state=="09")

colnames(race2015) <- c("state", "id", "total", "White", "Black", "Asian", "Hispanic")

race2015_per <- race2015
race2015_per$White <- round(race2015$White/race2015$total*100,2)
race2015_per$Black <- round(race2015$Black/race2015$total*100,2)
race2015_per$Asian <- round(race2015$Asian/race2015$total*100,2)
race2015_per$Hispanic <- round(race2015$Hispanic/race2015$total*100,2)

race2015_per <- gather(race2015_per, "Race", "per", 4:7) %>%
  select(Race, per)
```

```{r makeup, fig.width=11, fig.height=6}
gg <- ggplot(race_total_date, aes(x=mdy(DOWNLOAD.DATE), y=Percent)) + geom_bar(stat="identity") + facet_grid(.~Race) 
gg <- gg + geom_hline(data = race2015_per, aes(yintercept = per, color="red"))
gg <- gg + scale_color_discrete(name="Percent", labels="State resident population")
gg <- gg + labs(x=NULL, y="Percent", title=paste("Racial makeup of accused pre-trial inmates"),
                subtitle="Compared to overall demographic makeup of the state. Gaps in data indicate when the state did not update their daily feed",
                caption="SOURCE: Department of Corrections\nAndrew Ba Tran/TrendCT.org")
gg <- gg + theme_bw(base_family="Calibri")
gg <- gg + theme(text = element_text(size=16))
gg <- gg + theme(panel.border=element_blank())
gg <- gg + theme(plot.title=element_text(face="bold", family="Lato Black", size=22))
gg <- gg + theme(plot.subtitle=element_text(face="italic", size=9, margin=margin(b=12)))
gg <- gg + theme(plot.caption=element_text(size=12, margin=margin(t=10, r=80), color="#7a7d7e"))
gg <- gg + theme(plot.margin = unit(c(1, 5, 1, 10), "lines"))

gg
#ggsave(gg, file = "race_percent_time.png", width = 8, height = 6, type = "cairo-png")


## percent makeup compared to state stats


## per capita adjustment

```

```{r per_capita, fig.width=9, fig.height=6}

asian_r15 <- race2015$Asian
black_r15 <- race2015$Black
hispanic_r15 <- race2015$Hispanic
white_r15 <- race2015$White

race_total_date2 <- update %>%
  group_by(DOWNLOAD.DATE, RACE) %>%
  summarise(total=n()) %>%
  spread(RACE, total) %>%
  mutate(total = `AMER IND` + ASIAN + BLACK + HISPANIC + WHITE) %>%
  select(DOWNLOAD.DATE, ASIAN, BLACK, HISPANIC, WHITE) %>%
  mutate(ASIAN=ASIAN/asian_r15*10000, BLACK=BLACK/black_r15*10000, HISPANIC=HISPANIC/hispanic_r15*10000, WHITE=WHITE/white_r15*10000) %>%
  select(DOWNLOAD.DATE, ASIAN, BLACK, HISPANIC, WHITE) %>%
  gather("Race", "Per.Capita", 2:5) 

race_total_date2$Race <- str_to_title(race_total_date2$Race)


gg <- ggplot(race_total_date2, aes(x=mdy(DOWNLOAD.DATE), y=Per.Capita)) + geom_bar(stat="identity") + facet_grid(.~Race) 
gg <- gg + labs(x=NULL, y="Per 10,000 residents", title=paste("Racial makeup of accused pre-trial inmates."),
                subtitle="Normalized for demographic population.",
                caption="SOURCE: Department of Corrections\nAndrew Ba Tran/TrendCT.org")
gg <- gg + theme_bw(base_family="Calibri")
gg <- gg + theme(text = element_text(size=16))
gg <- gg + theme(panel.border=element_blank())
gg <- gg + theme(plot.title=element_text(face="bold", family="Lato Black", size=22))
gg <- gg + theme(plot.subtitle=element_text(face="italic", size=9, margin=margin(b=12)))
gg <- gg + theme(plot.caption=element_text(size=12, margin=margin(t=10, r=80), color="#7a7d7e"))
gg <- gg + theme(plot.margin = unit(c(1, 5, 1, 10), "lines"))

gg


```

```{r gender}
female_total <- update %>%
  group_by(DOWNLOAD.DATE, GENDER) %>%
  summarise(total=n()) %>%
  spread(GENDER, total) 

avg_females <- round(mean(female_total$F),0)
avg_males <- round(mean(female_total$M),0)

#round(avg_females/(avg_females+avg_males)*100)

median_females <- round(median(female_total$F),0)
median_males <- round(median(female_total$M),0)
```

Histogram of bail amounts by race

```{r histogram_bails, fig.width=9, fig.height=6}

bail_df <- update[!duplicated(update$IDENTIFIER), ]

#ggplot(data=bail_df, aes(BOND.AMOUNT)) + geom_histogram(bins = 500) + facet_grid(RACE~.)

#bail_df_upper <- subset(bail_df, BOND.AMOUNT >=100000)  
#bail_df_lower <- subset(bail_df, BOND.AMOUNT <100000)  

#ggplot(data=bail_df_upper, aes(BOND.AMOUNT)) + geom_histogram(bins = 500) + facet_grid(RACE~.)
#ggplot(data=bail_df_lower, aes(BOND.AMOUNT)) + geom_histogram(binwidth = 500) + facet_grid(RACE~.)


#range(bail_df$BOND.AMOUNT)

```

# Where pre-trial inmates are currently being held 
```{r where, fig.width=9, fig.height=6}

library(leaflet)
library(ggmap)

addresses <- read.csv("facilities.csv", stringsAsFactors=F)

bail_df <- update[!duplicated(update$IDENTIFIER), ]

bail_df$DOWNLOAD.DATE <- mdy(bail_df$DOWNLOAD.DATE)
bail_df <- arrange(bail_df, desc(DOWNLOAD.DATE))
last_day <- bail_df$DOWNLOAD.DATE[1]

last_day_df <- subset(update, mdy(DOWNLOAD.DATE)==last_day)

locations <- last_day_df %>%
  group_by(FACILITY) %>%
  summarise(total=sum(n())) %>%
  left_join(addresses)

# This function geocodes a location (find latitude and longitude) using the Google Maps API
geo <- geocode(location = locations$Address, output="latlon", source="google")

# add those coordinates to our dataset
locations$lon <- geo$lon
locations$lat <- geo$lat

m <- leaflet(locations) %>% addTiles('http://{s}.basemaps.cartocdn.com/dark_all/{z}/{x}/{y}.png', 
	attribution='Map tiles by <a href="http://stamen.com">Stamen Design</a>, <a href="http://creativecommons.org/licenses/by/3.0">CC BY 3.0</a> &mdash; Map data &copy; <a href="http://www.openstreetmap.org/copyright">OpenStreetMap</a>') 
m %>% setView(-72.690940, 41.651426, zoom = 8)
m %>% addCircles(~lon, ~lat, popup=locations$total, weight = 3, radius=locations$total/20*3.14, 
                 color="#ffa500", stroke = TRUE, fillOpacity = 0.8) 
m

```

### Men and women in jail awaiting trial

**Average number of women:** `r avg_females`

**Average number of men:** `r avg_males`

**Median number of women:** `r median_females`

**Median number of men:** `r median_males`

```{r females_race, fig.width=9, fig.height=6}
female_total <- update %>%
  filter(GENDER=="F") %>%
  group_by(DOWNLOAD.DATE, RACE) %>%
  summarise(total=n()) %>%
  group_by(RACE) %>%
  summarise(avg=mean(total))

gg <- ggplot(female_total, aes(x=RACE, y=avg)) + geom_bar(stat="identity") 
gg <- gg + labs(x=NULL, y="Average", title=paste("Racial makeup of female accused pre-trial inmates"),
                caption="SOURCE: Department of Correction\nAndrew Ba Tran/TrendCT.org")
gg <- gg + theme_bw(base_family="Calibri")
gg <- gg + theme(text = element_text(size=16))
#gg <- gg + theme(panel.grid.major=element_blank())
#gg <- gg + theme(panel.grid.minor=element_blank())
gg <- gg + theme(panel.border=element_blank())
#gg <- gg + theme(axis.ticks=element_blank())
#gg <- gg + theme(axis.text.x=element_blank())
gg <- gg + theme(plot.title=element_text(face="bold", family="Lato Black", size=22))
gg <- gg + theme(plot.subtitle=element_text(face="italic", size=9, margin=margin(b=12)))
gg <- gg + theme(plot.caption=element_text(size=12, margin=margin(t=10, r=80), color="#7a7d7e"))
gg <- gg + theme(plot.margin = unit(c(1, 1, 1, 1), "lines"))

gg
```

```{r males_race, fig.height=6, fig.width=9}
male_total <- update %>%
  filter(GENDER=="M") %>%
  group_by(DOWNLOAD.DATE, RACE) %>%
  summarise(total=n()) %>%
  group_by(RACE) %>%
  summarise(avg=mean(total))

gg <- ggplot(male_total, aes(x=RACE, y=avg)) + geom_bar(stat="identity") 
gg <- gg + labs(x=NULL, y="Average", title=paste("Racial makeup of male accused pre-trial inmates"),
                caption="SOURCE: Department of Correction\nAndrew Ba Tran/TrendCT.org")
gg <- gg + theme_bw(base_family="Calibri")
gg <- gg + theme(text = element_text(size=16))
#gg <- gg + theme(panel.grid.major=element_blank())
#gg <- gg + theme(panel.grid.minor=element_blank())
gg <- gg + theme(panel.border=element_blank())
#gg <- gg + theme(axis.ticks=element_blank())
#gg <- gg + theme(axis.text.x=element_blank())
gg <- gg + theme(plot.title=element_text(face="bold", family="Lato Black", size=22))
gg <- gg + theme(plot.subtitle=element_text(face="italic", size=9, margin=margin(b=12)))
gg <- gg + theme(plot.caption=element_text(size=12, margin=margin(t=10, r=80), color="#7a7d7e"))
gg <- gg + theme(plot.margin = unit(c(1, 1, 1, 1), "lines"))

gg
```

```{r avg_age}

avg_age <- update %>%
  group_by(RACE) %>%
  summarise(avg_age=round(mean(AGE),0), median_age=round(median(AGE),0))
```

### Average age by race

```{r avg_table}
kable(avg_age)
```

```{r facilities_total}
facilities <- update %>%
  group_by(DOWNLOAD.DATE, FACILITY) %>%
  summarise(total=n()) %>%
  group_by(FACILITY) %>%
  summarise(average=round(mean(total),0), median=round(median(total),0)) %>%
  arrange(-average)
```

### Average number of inmates by facility

```{r facilities_table}
kable(facilities)
```

```{r bonds_race}
# Bond amount total by race

bonds <- update %>%
  group_by(RACE) %>%
  summarise(avg_bond=mean(BOND.AMOUNT), median_bond=median(BOND.AMOUNT))
```

### Bond amount by race

```{r bonds_table}
kable(bonds)
```

```{r female_offense_list}
update$offense_pre <- gsub(", .*", "", update$OFFENSE)
update$offense_degree <- gsub(".*, ", "", update$OFFENSE)

update$offense_degree <- ifelse(update$offense_pre==update$offense_degree, "", update$offense_degree)

fem_offense_list <- update %>%
  filter(GENDER=="F") %>%
  group_by(DOWNLOAD.DATE, offense_pre) %>%
  summarise(count=n()) %>%
  group_by(offense_pre) %>%
  summarise(avg=round(mean(count),0)) %>%
  arrange(-avg) 
```

## Female offense list

```{r fem_of_table}
datatable(fem_offense_list)
```

```{r male_offense_list}
mal_offense_list <- update %>%
  filter(GENDER=="M") %>%
  group_by(DOWNLOAD.DATE, offense_pre) %>%
  summarise(count=n()) %>%
  group_by(offense_pre) %>%
  summarise(avg=round(mean(count),0)) %>%
  arrange(-avg) 
```

## Male offense list

```{r male_of_table}
datatable(mal_offense_list)
```

```{r male_offense_race}

mal_offense_race <- update %>%
  filter(GENDER=="M") %>%
  group_by(DOWNLOAD.DATE, offense_pre, RACE) %>%
  summarise(count=n()) %>%
  spread(RACE, count) %>%
  group_by(offense_pre) %>%
  summarise(amer.ind.avg=round(mean(`AMER IND`),0), asian.avg=round(mean(ASIAN),0), black.avg=round(mean(BLACK),0), hispanic.avg=round(mean(HISPANIC),0), white.avg=round(mean(WHITE),0)) %>%
  arrange(-white.avg) 
```

## Male offense list by race

```{r mor_table}
datatable(mal_offense_race)
```

```{r female_offense_race}

fem_offense_race <- update %>%
  filter(GENDER=="F") %>%
  group_by(DOWNLOAD.DATE, offense_pre, RACE) %>%
  summarise(count=n()) %>%
  spread(RACE, count) %>%
  group_by(offense_pre) %>%
  summarise(amer.ind.avg=round(mean(`AMER IND`),0), asian.avg=round(mean(ASIAN),0), black.avg=round(mean(BLACK),0), hispanic.avg=round(mean(HISPANIC),0), white.avg=round(mean(WHITE),0)) %>%
  arrange(-white.avg) 
```

## Female offense list by race

```{r fem_table}
datatable(fem_offense_race)
```

```{r percent_offense_table, warning=F, message=F, fig.height=11, fig.width=9}

mal_offense_race_percent <- update %>%
  filter(GENDER=="M") %>%
  group_by(DOWNLOAD.DATE, offense_pre, RACE) %>%
  summarise(count=n()) %>%
  spread(RACE, count) %>%
  group_by(offense_pre) %>%
  summarise(amer.ind.avg=round(mean(`AMER IND`),2), asian.avg=round(mean(ASIAN),0), black.avg=round(mean(BLACK),0), hispanic.avg=round(mean(HISPANIC),0), white.avg=round(mean(WHITE),0)) %>%
  arrange(-white.avg) %>%
  group_by(offense_pre) %>%
  mutate(total=sum(amer.ind.avg,asian.avg,black.avg,hispanic.avg,white.avg, na.rm=T)) %>%
  filter(total>20) %>%
  arrange(-total) %>%
  mutate(American.Indian.avg.per = round((amer.ind.avg/total)*100,0), Asian.avg.per = round(asian.avg/total*100,0),
Black.avg.per = round(black.avg/total*100,0), Hispanic.avg.per = round(hispanic.avg/total*100,0), White.avg.per = round(white.avg/total*100,0)) %>%
  select(offense_pre, total, American.Indian.avg.per, Asian.avg.per, Black.avg.per, Hispanic.avg.per, White.avg.per) %>%
  gather(Race, Percent, 3:7)

mal_offense_race_percent$Race <- gsub(".avg.per", "", mal_offense_race_percent$Race)
mal_offense_race_percent$Race <- gsub("\\.", " ", mal_offense_race_percent$Race)

#plotting it

library(plyr)
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

Data <- mal_offense_race_percent
Data[is.na(Data)] <- 0
Data <- ddply(Data, .(offense_pre), transform, pos=100-(cumsum(Percent)-(0.5*Percent)))


Data$offense_pre <- gsub(" AM", "", Data$offense_pre)
Data$offense_pre <- gsub(" DF", "", Data$offense_pre)
Data$offense_pre <- gsub(" AF", "", Data$offense_pre)
Data$offense_pre <- gsub(" CF", "", Data$offense_pre)
Data$offense_pre <- gsub("  F", "", Data$offense_pre)
Data$offense_pre <- str_trim(Data$offense_pre)

Data[Data == 0] <- NA
Data <- Data[ order(-Data$total), ]

Data$offense_pre <- factor(Data$offense_pre, levels = Data$offense_pre[order(Data$total)])

#Data$Race <- factor(Data$Race)


#Data$Race <- factor(Data$Race,levels(Data$Race)[c(6,5,4,3,2,1)])


gg <- ggplot(Data, aes(x = offense_pre, y = Percent)) 
gg <- gg + geom_bar(aes(fill = Race, order=-as.numeric(Race)), stat="identity") 
gg <- gg + geom_text(aes(label = Percent, y = pos), size = 4) 
gg <- gg + scale_fill_manual(values=cbPalette)
gg <- gg + coord_flip() 
gg <- gg + geom_rect(data=Data, aes(ymin=102, ymax=114, xmin=-Inf, xmax=Inf), fill="#efefe3")
gg <- gg + geom_text(data=Data, aes(label=round(total,0), x=offense_pre, y=108), fontface="bold", size=4, family="Calibri")
gg <- gg + labs(x=NULL, y="Percent", title="Offenses that accused pre-trial inmates face by race",
                subtitle="Percent of race and total facing offenses. Based on rolling average of inmates per day.",
                caption="Department of Correction\nAndrew Ba Tran/TrendCT.org")
gg <- gg + theme_bw(base_family="Calibri")
gg <- gg + theme(panel.grid.major=element_blank())
gg <- gg + theme(panel.grid.minor=element_blank())
gg <- gg + theme(panel.border=element_blank())
gg <- gg + theme(axis.ticks=element_blank())
gg <- gg + theme(axis.text.x=element_blank())
#gg <- gg + theme(text = element_text(size=20))
gg <- gg + theme(plot.title=element_text(face="bold", family="Lato Black", size=16))
gg <- gg + theme(plot.subtitle=element_text(face="italic", size=10,margin=margin(b=12)))
gg <- gg + theme(plot.caption=element_text(size=10, margin=margin(t=16), color="#7a7d7e"))
gg
```


```{r bond_avg_race, warning=F, message=F}
detach("package:plyr", unload=TRUE)

mal_offense_race_bond <- update %>%
  filter(GENDER=="M") %>%
  group_by(DOWNLOAD.DATE, OFFENSE, RACE) %>%
  summarise(total=n(),avg.bond=mean(BOND.AMOUNT)) %>%
  select(DOWNLOAD.DATE, OFFENSE, total, RACE, avg.bond) %>%
  spread(RACE, avg.bond) %>%
  group_by(OFFENSE) %>%
  summarise(total = sum(total), amer.ind.avg=round(mean(`AMER IND`, na.rm=T),2), asian.avg=round(mean(ASIAN, na.rm=T),0), black.avg=round(mean(BLACK, na.rm=T),0), hispanic.avg=round(mean(HISPANIC, na.rm=T),0), white.avg=round(mean(WHITE, na.rm=T),0)) %>%
  arrange(-white.avg) %>%
  filter(total>20)

mal_offense_race_bond$what <- ifelse(((mal_offense_race_bond$white.avg < mal_offense_race_bond$black.avg) & (mal_offense_race_bond$white.avg < mal_offense_race_bond$hispanic.avg)), "white", "minority")

```

## Bond amount

```{r bond_amount_table}
datatable(mal_offense_race_bond)
```


```{r time_in_jail}

days_list <- unique(update$DOWNLOAD.DATE)
update$auth.date <- mdy(update$DOWNLOAD.DATE)
update$first.date <- ifelse(mdy(update$LATEST.ADMISSION.DATE)>mdy("07/01/2016"), as.character(mdy(update$LATEST.ADMISSION.DATE)), as.character(mdy("07/01/2016"))) 
update$last.date <- ""

#3427

for (i in 2:length(days_list)) {
  date <- days_list[i-1]
  update_sub <- subset(update, DOWNLOAD.DATE==date)
  update_sub_solo <- update_sub[c("DOWNLOAD.DATE", "IDENTIFIER", "first.date")]
  
  date2 <- days_list[i]
  update_sub2 <- subset(update, DOWNLOAD.DATE==date2)
  update_sub_solo2 <- update_sub2[c("DOWNLOAD.DATE", "IDENTIFIER", "first.date")]
  update_sub_combined <- full_join(update_sub_solo, update_sub_solo2, by="IDENTIFIER")

  update_sub_combined$first.date <- ifelse(is.na(update_sub_combined$DOWNLOAD.DATE.x), as.character(mdy(as.character(update_sub_combined$DOWNLOAD.DATE.y))), as.character(ymd(update_sub_combined$first.date.x)))
  update_sub_combined$last.date <- ifelse(is.na(update_sub_combined$DOWNLOAD.DATE.y), as.character(mdy(as.character(update_sub_combined$DOWNLOAD.DATE.x))), "")
  
  update_sub_combined <- update_sub_combined[c("IDENTIFIER","first.date","last.date")]
  
  if (i == 2) {
    update_all <- update_sub_combined
  } else {
    update_all <- rbind(update_all, update_sub_combined)
  }
  update_all_complete <- update_all[!duplicated(update_all), ]
  
}

dupes <- update_all_complete[duplicated(update_all_complete$IDENTIFIER),]
dupes$dupe <- "duplicate"
dupes <- dupes[c("IDENTIFIER", "dupe")]

update_all_complete <- left_join(update_all_complete, dupes)

duped <- subset(update_all_complete, dupe=="duplicate")
duped <- arrange(duped, IDENTIFIER)

for (i in 1:nrow(duped)) {
  if ((duped$IDENTIFIER[i]== duped$IDENTIFIER[i+1]) & (i !=nrow(duped))) {
    
    if (duped$first.date[i] == duped$first.date[i+1]) {
      duped$last.date[i] <- ifelse(duped$last.date[i]=="", duped$last.date[i+1], duped$last.date[i])
      duped$last.date[i+1] <- ifelse(duped$last.date[i]=="", duped$last.date[i+1], duped$last.date[i])
      
    }
    
    if (duped$first.date[i] > duped$first.date[i+1]) {
      duped$first.date[i] <- duped$first.date[i+1]
    }
    
    if (duped$first.date[i] < duped$first.date[i+1]) {
      duped$first.date[i + 1] <- duped$first.date[i]
    }
    
  }
  
}

duped <- unique(duped)

update_all_complete_final <- subset(update_all_complete, is.na(dupe))
update_all_complete_final <- rbind(update_all_complete_final, duped)

update$auth.date <- NULL
update$first.date <- NULL
update$last.date <- NULL

update <- left_join(update, update_all_complete_final, by="IDENTIFIER")
update$dupe <- NULL
update$first.date[update$first.date == "2016-07-01"] <- NA
update$total.days <- ymd(update$last.date) - ymd(update$first.date)

write.csv(table(update$total.days, update$RACE), "race_days.csv")
```

## Median days spent in jail (of those who came and left between July 1 and today

```{r time_table}
median_d <- update %>%
  group_by(RACE) %>%
  summarise(Days=median(total.days, na.rm=T))

kable(median_d)
```

```{r avg_time_table}
avg_d <- update %>%
  group_by(RACE) %>%
  summarise(Days=mean(total.days, na.rm=T))
kable(median_d)

out <- subset(update, !is.na(total.days))
still_in <- subset(update, is.na(total.days))
out_yet <- subset(still_in, last.date=="")
still_in <- subset(still_in, last.date!="")


out2 <- count(out[unique(out$IDENTIFIER),])
out_yet2 <- count(out_yet[unique(out_yet$IDENTIFIER),])
still_in2 <- count(still_in[unique(still_in$IDENTIFIER),])
```

## Inmate turnover

Pre-trial inmates who arrived before July 1 and has since left jail
* `r out_yet2`

Pre-trial inmates who arrived after July 1 and has since left jail
* `r out2`

Pre-trial inmates who are still in jail
* `r still_in2`

-----

## Inmates still in jail by race

```{r total_inmates1}
total_inmates <- length(unique(update$IDENTIFIER))


still_in_t <- still_in[unique(still_in$IDENTIFIER),] %>%
  group_by(RACE) %>%
  summarise(count=n())

kable(still_in_t)
```

## Inmates still who came in to jail and have left

```{r in_out}

out_t <- out[unique(out$IDENTIFIER),] %>%
  group_by(RACE) %>%
  summarise(count=n())

kable(out_t)

```

## Inmates who came in before July 1 and have since left

```{r out_yet}
out_yet_t <- out_yet[unique(out_yet$IDENTIFIER),] %>%
  group_by(RACE) %>%
  summarise(count=n())

kable(out_yet_t)
```